---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# Replication code and data for: "The heterogeneous relationship between income and inequality: a panel co-integration approach"

[Svenja Flechtern](http://www.svenjaflechtner.de/) and 
[Claudius Gräbner](https://claudius-graebner.com/)

Published as: Flechtner, S. and Gräbner, C. (2019): 
The heterogeneous relationship between income and inequality:
a panel co-integration approach, in: *Economics Bulletin*, Vol. 39(4), p. 2540-2549.
[Available online](http://www.accessecon.com/Pubs/EB/2019/Volume39/EB-19-V39-I4-P237.pdf) (open access).

## Description of code

The dataset used in the paper gets recreated from the raw data as it can be 
downloaded from the web via `R/create_dataset.R` (see below). 

The estimation results can be replicated via the Stata code in `stata/do-file_Flechtner_Graebner_2019_gdp-ineq-cointegration-short`.

The visualizations and the country classification table can be replicated using
the R code in `R/cointegration-results-analysis.R`.

The Stata code has been written by Svenja Flechtner, the R code by Claudius Gräbner.

All code assumes the working directory to be organized as in this Github repo.

## Description of results

The file `output/cointegration_results` summarizes the results in a readable way.
The csv version is used as an input for the visualization in the paper.

It contains the following variables:

| Column  | Description                                                          | 
| --------|:--------------------------------------------------------------------:| 
| country | country name abbreviation as in dataset |
| id  | id created during analysis (see stata do-file) |
| gini_mkt_coeff | coefficient obtained from PDOLS estimator for market gini; see stata do-file. recovered from stata log file `stata/logfile_ginimarketln-GDPpercapitaln` |
| gini_disp_coeff | coefficient obtained from PDOLS estimator for disposable gini; see stata do-file. recovered from stata log file `stata/logfile_ginidispln-GDPpercapitaln` |
| gini_mkt_t | t statistic for coefficient obtained from PDOLS estimator for market gini; see stata do-file. recovered from stata log file `output/logfile_ginimarketln-GDPpercapitaln` |
| gini_disp_t | t statistic for coefficient obtained from PDOLS estimator for disposable gini; see stata do-file. recovered from stata log file `stata/logfile_ginidispln-GDPpercapitaln` |
| gini_mkt_t_abs | absolute value of `gini_mkt_t` |
| gini_disp_t_abs | absolute value of `gini_disp_t` |
| no_obs | number of observations |
| df | degrees of freedom |
| critical_value | Critical values of Student's t distribution with respective number of degrees of freedom for `alpha=0.975` |
| gini_mkt_sign5 | dummy variable with 1 if `gini_mkt_t_abs` >= critival_value, 0 otherwise |
| gini_disp_sign5 | dummy variable with 1 if `gini_disp_t_abs` >= critival_value, 0 otherwise |

## Description of data

The dataset gets recreated via `R/create_dataset.R`. Here we assume you have dowloaded
the [Penn World Table](https://www.rug.nl/ggdc/productivity/pwt/) 9.1 
(downloaded September 17, 2019, named `data/pwt91.dta`)
and the [Standardized World Income Inequality Database](https://fsolt.org/swiid/) 
(downloaded September 17, 2019, named `data/swiid.dta`).

| Variable| Description             | Unit                   | Source    |
| --------|:-----------------------:| :---------------------:|:---------:|
| country | The iso3 country code   | iso3c                  | [Info][1] |
| year    | The year of observation | year                   | NA |
| gdp_pc_chppp | Per-capita income (expenditure-side) |  Real GDP at chained PPP | [Penn World tables 9.1][2] |
| gini_solt_disp  | Post-tax inequality  | Gini index        | [SWIID Version 8.1][3] |
| gini_solt_mkt   | Pre-tax inequality   | Gini index        | [SWIID Version 8.1][3] |

[1]: https://unstats.un.org/unsd/tradekb/knowledgebase/country-code
[2]: https://www.rug.nl/ggdc/productivity/pwt/
[3]: https://doi.org/10.7910/DVN/LM4OWF

The dataset relevant to reproduce the results is `data/dataset.csv`.
The file `data/dataset_raw.csv` is created directly from the raw data and only
excludes countries with less than 26 observations. I still contains those without 
a unit root in their Gini time series. As explained in the paper, these countries 
have been removed before the analysis and they are not contained in the file 
`data/dataset.csv` any more. See the file `R/create_dataset.R` for details, and
the file `stata/do-file_Flechtner_Graebner_2019_gdp-ineq-cointegration-short_country_unitroots` for the respective unit root tests.

### Descriptive statistics on the sample level

```{r, echo=FALSE, message=FALSE}
library(countrycode)
library(data.table)
library(here)
library(knitr)
library(tidyverse)
full_data <- fread(here("data/dataset.csv"))
descriptives_agg <- full_data %>%
  group_by() %>%
  summarise(`First observation`=min(year),
            `Last observation`=max(year),
            `Total number of observations`=n(), 
            `GDP per capita (mean)`=mean(gdp_pc_chppp),
            `GDP per capita (sd)`=sd(gdp_pc_chppp),
            `Post-tax gini (mean)`=mean(gini_disp),
            `Post-tax gini (sd)`=sd(gini_disp),
            `Pre-tax gini (mean)`=mean(gini_mkt),
            `Pre-tax gini (sd)`=sd(gini_mkt)) %>%
  ungroup() %>%
  gather()
kable(descriptives_agg, 
      caption = "Descriptive statistics for the data used (aggregated)", 
      digits = 2)
```


### Descriptive statistics on the country level

```{r, echo=FALSE, message=FALSE}
descriptives <- full_data %>%
  group_by(country) %>%
  summarise(y_min=min(year),
            y_max=max(year),
            n=n(), 
            GDPpc_mean=mean(gdp_pc_chppp),
            GDPpc_sd=sd(gdp_pc_chppp),
            GINIdisp_mean=mean(gini_disp),
            GINIdisp_sd=sd(gini_disp),
            GINImkt_mean=mean(gini_mkt),
            GINImkt_sd=sd(gini_mkt)) %>%
  ungroup() %>%
  unite("period", y_min, y_max, sep = "-") %>%
  mutate(country=countrycode(country, "iso3c", "country.name"))
kable(descriptives, 
      caption = "Descriptive statistics for the data used (country level)", 
      digits = 2)
```


### Citations for the data

Feenstra, Robert C., Robert Inklaar and Marcel P. Timmer (2015): 
The Next Generation of the Penn World Table, *American Economic Review*, Vol. 105(10), 
3150-3182, available for download at [www.ggdc.net/pwt](www.ggdc.net/pwt)

Solt, Frederick (2019): *Measuring Income Inequality Across Countries and Over Time: The Standardized World Income Inequality Database*, SWIID Version 8.1, May 2019. 
